\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}

### 2014-07-02 paper/subdocuments/methods.Rnw
added subsubtitel methods.Rnw

@@

% Probability Density function as a base for the RD
The goal of RD is to study the differences between two distributions. A common example could be the income distribution for man and women. Subject of the comparison can also be two distributions describing the same population, but stemming from two different data sources like survey or tax data or even comparison of the same source/population but for different time points, like the income distribution out of tax data for Switzerland today compared to an earlier time point. To describe what we are going to compare, we start with defining the two distributions. One represents the reference population $Y_{0}$ and the other the comparison population $Y$. $Y$ represents our measure of interest (income). A first visual approach is to compare the two probability density functions (PDF). The PDF is a function $f(y)$ which describes the distribution of probability over the outcome set and is defined for all possible values of $y$. Also this function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%. Out of the comparison of the PDF, it is possible to see, which values of $x$ are more and less probable. This allows already to spot distributional differences over the whole scale of $Y$ visually. \\

The PDF can be characterized by its cumulative distribution function (CDF). The CDF can be formulated as $F(y)$, which represents the cumulative probability at point $x$, which is the probability that a randomly chosen value is less than or equal to $y$. The relative distribution of $Y$ to $Y_{0}$ is then defined as

@@

We can calculate the Probability Density Function $g(r)$ of RD, where $r$ represents the proportion of values and $F_{0}^{-1}(r)$ is the inverse cumulative distribution function, also called the quantile function. $g(r)$ can be interpreted as a density ratio, which is defined as the ratio of these two quantities evaluated at every percentile of the reference distribution [0,1]. With a complete overlap of both distributions the probability density function of the RD is 1 at every point of the PDF. On the other hand, values higher than 1 represent higher probabilities in the comparison distribution than in the references distribution at this specific point and values lower than 1 respectively represent lower probabilities. It is a proper PDF in the sense that it integrates to 1 over the unit interval.\\

@@

% RD based Summary Meausres
Distributional polarization is of particular interest in the study of inequality. However, common inequality indicators (for example Gini or Theil’s index) are not designed to distinguish between growth in the upper and lower tails. Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail) or downgrading (increases in lower tail). The polarization index plays the same role as the difference in Gini-indices, but it can be decomposed to compare the growth in the upper and lower tails. Because it is based on the relative distribution it provides a simple link between what is observed in the graphical display and what is measured by the numerical summary. \\

>>>>

\subsubsection{Relative distribution (RD) methods }

@@

% Probability Density function as a base for the RD
The goal of RD is to study the differences between two distributions. A common example could be the income distribution for man and women. Subject of the comparison can also be two distributions describing the same population, but stemming from two different data sources like survey or tax data or even comparison of the same source/population but for different time points, like the income distribution out of tax data for Switzerland today compared to an earlier time point. To describe what we are going to compare, we start with defining the two distributions. One represents the reference population $Y_{0}$ and the other the comparison population $Y$. $Y$ represents our measure of interest (income). A first visual approach is to compare the two probability density functions (PDF). The PDF is a function $f(y)$ which describes the distribution of probability over the outcome set and is defined for all possible values of $y$. Also this function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%. Out of the comparison of the PDF, it is possible to see, which values of $y$ are more and less probable. This allows already to spot distributional differences over the whole scale of $Y$ visually. \\

The PDF can be characterized by its cumulative distribution function (CDF). The CDF can be formulated as $F(y)$, which represents the cumulative probability at point $y$, which is the probability that a randomly chosen value is less than or equal to $y$. The relative distribution of $Y$ to $Y_{0}$ is then defined as

@@

We can calculate the Probability Density Function $g(r)$ of R, where $r$ represents the proportion of values and $F_{0}^{-1}(r)$ is the inverse cumulative distribution function, also called the quantile function. $g(r)$ can be interpreted as a density ratio, which is defined as the ratio of these two quantities evaluated at every percentile of the reference distribution [0,1]. With a complete overlap of both distributions the probability density function of the R is 1 at every point of the PDF. On the other hand, values higher than 1 represent higher probabilities in the comparison distribution than in the references distribution at this specific point and values lower than 1 respectively represent lower probabilities. It is a proper PDF in the sense that it integrates to 1 over the unit interval.\\

@@

% R based Summary Meausres
Distributional polarization is of particular interest in the study of inequality. However, common inequality indicators (for example Gini or Theil’s index) are not designed to distinguish between growth in the upper and lower tails. Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail) or downgrading (increases in lower tail). The polarization index plays the same role as the difference in Gini-indices, but it can be decomposed to compare the growth in the upper and lower tails. Because it is based on the relative distribution it provides a simple link between what is observed in the graphical display and what is measured by the numerical summary. \\
